---
title: "Predicting interactions"
---


```{r inlude= FALSE}
# load packages
library(tidyverse)
library(ade4)
library(gbm)
library(randomForest)
library(ape)
library(cluster)
library(reshape)
library(ecodist)
library(knitr)
library(kableExtra)

# in use
library(igraph)
library(ggpubr)
library(magrittr)
library(stringr) # strings
library(reshape2) # melt
library(varhandle) # to dummy

source("./functions/functions_metawebs.R")
source("./functions/functions_plotting.R")
```

## Step 1:Building randomforest models
```{r}
# load random forest based on only positive interactions
load("./output/rf_mod_diff_observed_metaweb.RData")

# load random forest based on IB
load("./output/rf_mod_ib_filter2.RData")
load("./output/hybrid_rf_mod_ib.RData")

rf_object <- rf_mod_diff_observed_metaweb

# load ib interactions for prediction
load(file = "./data/hybrid_interact_con_res_ib.RData")
load(file = "./data/hybrid_interact_con_res_ib_with_species_codes.RData")

# load traits
load("./data/hybrid_data_trait_phylo.RData")

hybrid_interact_con_res_ib <-
  hybrid_interact_con_res_ib %>% drop_na(.)

hybrid_interact_con_res_ib_with_species_codes <- hybrid_interact_con_res_ib_with_species_codes %>% drop_na(.)
```

## Step 2: Mean decrease Gini (contribution of each trait)
The ‘mean decrease Gini’ (total decrease in node impurities from splitting on the variable) is proportional to the contribution of each trait in the model. High mean decrease Gini represents a high importance of the variable. 
```{r fig.cap="trait contribution randomForest", fig.width=7}
# extract contribution of each trait
importance_traits <- importance(rf_object, type = 2)
table <- data.frame(traits = row.names(importance_traits), importance_traits, row.names = NULL)
table_brose <- data.frame(table[order(table$IncNodePurity, decreasing = T),], row.names = NULL)
#
## extract contribution of each trait IberianPonds
importance_traits <- importance(hybrid_rf_mod_ib, type = 2)
table <- data.frame(traits = row.names(importance_traits), importance_traits, row.names = NULL)
table_ib <- data.frame(table[order(table$IncNodePurity, decreasing = T),], row.names = NULL)
```

#Gini plots
```{r}
png(file="./figures/giniplots_hybrid.png", width = 10, height = 6, units = 'in', res= 300)
p1<- table_brose %>%
ggplot(aes(x=reorder(traits, IncNodePurity), y=IncNodePurity)) +
 geom_boxplot()+
 coord_flip()+ ylab("trait contribution") + theme_bw()+ theme(axis.title.y=element_blank(), text=element_text(size=11))+ ggtitle("Training dataset")
#+ ylim(0, 6000)


p2<- table_ib %>%
ggplot(aes(reorder(traits, IncNodePurity), y=IncNodePurity)) +
 geom_boxplot()+
 coord_flip()+ ylab("trait contribution") + theme_bw()+ theme(axis.title.y=element_blank(),
   text=element_text(size=11))+ ggtitle("Iberian Pond Network")
#+ylim(0, 15000)

ggarrange(p1, p2)
dev.off()
```


### Step 3: Predicting interaction in the Iberian Pond Network (IPN) based on Brose dataset (independent interactions dataset) randomforest model
```{r, fig.cap = "Predicted IberianPonds (using randomForest Brose)"}
set.seed(1234)

# run predict function with interact_con_res_ib as "newdata"
hybrid_interact_con_res_ib$pred1 <- predict(rf_object, newdata = hybrid_interact_con_res_ib, predict.all = FALSE)

# attach species codes to dataframe
hybrid_interact_con_res_ib <- cbind(hybrid_interact_con_res_ib_with_species_codes[, 1:2], hybrid_interact_con_res_ib)
```


```{r}
predator_interactions <- hybrid_interact_con_res_ib %>%
  dplyr::filter(con_feeding_type.predator == "1")
# omnivores
omnivores_interactions <- hybrid_interact_con_res_ib %>%
  dplyr::filter(con_feeding_type.omnivore == 1) # here we have others (herbivore, omnivore, detritivore) except predators
herbivore_interactions <- hybrid_interact_con_res_ib %>%
  dplyr::filter(con_feeding_type.herbivore == "1") # here we have only herbivores

#getting top 25% of the interactions for different interaction types
T_pred <- quantile(predator_interactions$pred1, c(.75))
T_omni <- quantile(omnivores_interactions$pred1, c(.75))
T_herb <- quantile(herbivore_interactions$pred1, c(.75))

hybrid_rf_interactions <- hybrid_interact_con_res_ib

### select predator interactions
hybrid_rf_interactions[hybrid_rf_interactions$con_feeding_type.predator==1 & hybrid_rf_interactions$pred1 >= T_pred,]$value <- 1
hybrid_rf_interactions[hybrid_rf_interactions$con_feeding_type.predator==1 & hybrid_rf_interactions$pred1 < T_pred,]$value <- 0

### select omnivore interactions
hybrid_rf_interactions[hybrid_rf_interactions$con_feeding_type.omnivore==1 & hybrid_rf_interactions$pred1 >= T_omni,]$value <- 1
hybrid_rf_interactions[hybrid_rf_interactions$con_feeding_type.omnivore==1 & hybrid_rf_interactions$pred1 < T_omni,]$value <- 0
  
### select herbivore interactions
hybrid_rf_interactions[hybrid_rf_interactions$con_feeding_type.herbivore==1 & hybrid_rf_interactions$pred1 >= T_herb,]$value <- 1
hybrid_rf_interactions[hybrid_rf_interactions$con_feeding_type.herbivore==1 & hybrid_rf_interactions$pred1 < T_herb,]$value <- 0

colnames(hybrid_rf_interactions)[colnames(hybrid_rf_interactions)=="value"] <- "weight"

hybrid_rf_edges <-
  hybrid_rf_interactions %>%
  dplyr::select(
    con_code, res_code, weight
  )
#rename and change the order
names(hybrid_rf_edges)[names(hybrid_rf_edges) == "con_code"] <- "from"
names(hybrid_rf_edges)[names(hybrid_rf_edges) == "res_code"] <- "to"
hybrid_rf_graph <- prepare_graphs_plotting(hybrid_rf_edges, hybrid_data_trait_phylo, .column_name = "unique_id")

# add detritus
hybrid_rf_graph <- add_detritus_to_graph(.graph = hybrid_rf_graph)


# save prediction objects with different consensus filters (RF only)
save(hybrid_rf_interactions, file = "./output/hybrid_rf_interactions.RData")
save(hybrid_rf_edges, file = "./output/hybrid_rf_edges.RData")
save(hybrid_rf_graph, file = "./output/hybrid_rf_graph.RData")
```
