---
title: "Trait-matching model"
---

## Description:
This script uses one independent dataset (GATEWay database; Brose, 2018, Brose et al., 2019) to infer food web interactions in Iberian Ponds hybrid dataset (morphological + eDNA) (Pereira et al.2021), by applying trait-matching (adapted from Laigle et al. 2018).


```{r include = FALSE}
# load packages
library(randomForest)
library(plyr)
library(tidyr)
```

```{r include=FALSE}
# load data
load(file = "./data/hybrid_interact_con_res_ib.RData")
```

```{r include=FALSE}
# remove links with missing biomass
hybrid_interact_con_res_ib <- 
  hybrid_interact_con_res_ib %>% drop_na(.)
```

# Building randomforest models for IPN Hybrid
```{r}
# run random forest 
hybrid_rf_mod_ib <- randomForest(value ~ ., data = hybrid_interact_con_res_ib, importance=TRUE) 
# if the code breaks at this step due to memory issues, the randomForest object that is generated here is also provided in "./output/hybrid_rf_mod_ib.RData".

# save objects
#save(hybrid_rf_mod_ib, file = "./output/hybrid_rf_mod_ib.RData")
```


## References
Brose U. 2018. GlobAL daTabasE of traits and food web Architecture (GATEWAy) v.1.0. iDiv Data Repository accessed 17 April 2019.
Brose U, Archambault P, Barnes AD, Bersier L-F, Boy T, Canning-Clode J, Conti E, Dias M, Digel C, Dissanayake A, Flores AAV, Fussmann K, Gauzens B, Gray C, Häussler J, Hirt MR, Jacob U, Jochum M, Kéfi S, McLaughlin O, MacPherson MM, Latz E, Layer-Dobra K, Legagneux P, Li Y, Madeira C, Martinez ND, Mendonça V, Mulder C, Navarrete SA, O’Gorman EJ, Ott D, Paula J, Perkins D, Piechnik D, Pokrovsky I, Raffaelli D, Rall BC, Rosenbaum B, Ryser R, Silva A, Sohlström EH, Sokolova N, Thompson MSA, Thompson RM, Vermandele F, Vinagre C, Wang S, Wefer JM, Williams RJ, Wieters E, Woodward G, Iles AC. 2019. Predator traits determine food-web architecture across ecosystems. Nature Ecology & Evolution 3: 919–927.
Laigle I, Aubin I, Digel C, Brose U, Boulangeat I, Gravel D. 2018. Species traits as drivers of food web structure. Oikos 127: 316–326.
Pereira CL, Gilbert MTP, Araújo MB, Matias MG. 2021. Fine‐tuning biodiversity assessments: A framework to pair eDNA metabarcoding and morphological approaches. Methods in Ecology and Evolution 12: 2397–2409.