---
title: "R Notebook"
---

## Description:
This script uses one independent dataset (Brose et al. 2019) to infer food web interactions in Iberian Ponds hybrid dataset (morphological+ edna), by applying trait matching (adapted from Laigle et al. 2018).

```{r inlude= FALSE}
# load packages
library(tidyverse)
library(ade4)
library(gbm)
library(randomForest)
library(ape)
library(cluster)
library(reshape)
library(ecodist)
library(knitr)
library(kableExtra)

# in use
library(igraph)
library(ggpubr)
library(magrittr)
library(stringr) # strings
library(reshape2) # melt
library(varhandle) # to dummy

source("./functions/functions_metawebs.R")
source("./functions/functions_plotting.R")
```

## Step 4: Building randomforest models
### Step 4a: randomforest model of brose dataset
```{r}
# load random forest based on only positive interactions
load("./output/rf_mod_diff_observed_metaweb.RData")

# load random forest based on IB
load("./output/rf_mod_ib_filter2.RData")
load("./output/hybrid_rf_mod_ib.RData")

rf_object <- rf_mod_diff_observed_metaweb

# load ib interactions for prediction
load(file = "./data/hybrid_interact_con_res_ib.RData")
load(file = "./data/hybrid_interact_con_res_ib_with_species_codes.RData")

# load traits
load("./data/hybrid_data_trait_phylo.RData")

hybrid_interact_con_res_ib <-
  hybrid_interact_con_res_ib %>% drop_na(.)

hybrid_interact_con_res_ib_with_species_codes <-
  hybrid_interact_con_res_ib_with_species_codes %>% drop_na(.)
```

### Step 5 Mean decrease Gini (contribution of each trait)
The ‘mean decrease Gini’ (total decrease in node impurities from splitting on the variable) is proportional to the contribution of each trait in the model. High mean decrease Gini represents a high importance of the variable. 

```{r fig.cap="trait contribution randomForest", fig.width=7}
# extract contribution of each trait
importance_traits <- importance(rf_object, type = 2)
table <- data.frame(traits = row.names(importance_traits), importance_traits, row.names = NULL)
table_brose <- data.frame(table[order(table$IncNodePurity, decreasing = T),], row.names = NULL)
#
## extract contribution of each trait IberianPonds
importance_traits <- importance(hybrid_rf_mod_ib, type = 2)
table <- data.frame(traits = row.names(importance_traits), importance_traits, row.names = NULL)
table_ib <- data.frame(table[order(table$IncNodePurity, decreasing = T),], row.names = NULL)
```


```{r}
png(file="./figures/giniplots_hybrid.png", width = 10, height = 6, units = 'in', res= 300)
p1<- table_brose %>%
ggplot(aes(x=reorder(traits, IncNodePurity), y=IncNodePurity)) +
 geom_boxplot()+
 coord_flip()+ ylab("trait contribution") + theme_bw()+ theme(axis.title.y=element_blank(), text=element_text(size=11))+ ggtitle("Training dataset")
#+ ylim(0, 6000)


p2<- table_ib %>%
ggplot(aes(reorder(traits, IncNodePurity), y=IncNodePurity)) +
 geom_boxplot()+
 coord_flip()+ ylab("trait contribution") + theme_bw()+ theme(axis.title.y=element_blank(),
   text=element_text(size=11))+ ggtitle("Iberian Pond Network")
#+ylim(0, 15000)

ggarrange(p1, p2)
dev.off()
```


### Step 6a: Predicting interaction in the IberianPonds based on Brose's randomforest model
```{r, fig.cap = "Predicted IberianPonds (using randomForest Brose)"}
set.seed(1234)

# run predict function with interact_con_res_ib as "newdata"
hybrid_interact_con_res_ib$pred1 <- predict(rf_object, newdata = hybrid_interact_con_res_ib, predict.all = FALSE)

# attach species codes to dataframe
hybrid_interact_con_res_ib <- cbind(hybrid_interact_con_res_ib_with_species_codes[, 1:2], hybrid_interact_con_res_ib)
```


# Calculate TSS for different interactions (predator, grazer, herbivore)
```{r}
TR_all <- 0.2

# all interactions
# plot interactions probabilities

TSS <- CalcTSS(hybrid_interact_con_res_ib, thresh = TR_all)

# Predators
# plot interactions probabilities
predator_interactions <- hybrid_interact_con_res_ib %>%
  dplyr::filter(con_feeding_type.predator == "1")

TSS_pred <- CalcTSS(predator_interactions, thresh = quantile(predator_interactions$pred1, c(.75)))

# omnivores
omnivores_interactions <- hybrid_interact_con_res_ib %>%
  dplyr::filter(con_feeding_type.omnivore == 1) # here we have others (herbivore, omnivore, detritivore) except predators

TSS_omni <- CalcTSS(omnivores_interactions, thresh = quantile(omnivores_interactions$pred1, c(.75)))

# Herbivores
herbivore_interactions <- hybrid_interact_con_res_ib %>%
  dplyr::filter(con_feeding_type.herbivore == "1") # here we have only herbivores

TSS_herb <- CalcTSS(herbivore_interactions, thresh = quantile(herbivore_interactions$pred1, c(.75)))

# summary figure
prob_distribs <- rbind(
  cbind(filter = rep("all links", length(hybrid_interact_con_res_ib$pred1)), prob = hybrid_interact_con_res_ib$pred1),
  cbind(filter = rep("predation", length(predator_interactions$pred1)), prob = predator_interactions$pred1),
  cbind(filter = rep("omnivory", length(omnivores_interactions$pred1)), prob = omnivores_interactions$pred1),
  cbind(filter = rep("herbivory", length(herbivore_interactions$pred1)), prob = herbivore_interactions$pred1)
) %>%
  as.data.frame() %>%
  mutate(
    prob = as.numeric(prob),
    filter = factor(filter, levels = c("all links", "predation", "omnivory", "herbivory"))
  )

plot_hybrid_distribs <-
  prob_distribs %>%
  ggplot(aes(x = prob)) +
  facet_grid(vars(filter)) +
  geom_histogram(aes(fill = filter))

save(plot_hybrid_distribs, file = "./figures/plot_hybrid_distribs.RData")

# summary table
tss_full <-
  cbind(
    data.frame(thresh_levels = c(TR_all, quantile(predator_interactions$pred1, c(.75)), quantile(omnivores_interactions$pred1, c(.75)), quantile(herbivore_interactions$pred1, c(.75)))),
    rbind(
      TSS = round(cbind(TSS[, 1:4] / sum(TSS[, 1:4]) * 100, TSS = TSS[, 5]), 2),
      TSS_pred = round(cbind(TSS_pred[, 1:4] / sum(TSS_pred[, 1:4]) * 100, TSS = TSS_pred[, 5]), 2),
      TSS_omni = round(cbind(TSS_omni[, 1:4] / sum(TSS_omni[, 1:4]) * 100, TSS = TSS_omni[, 5]), 2),
      TSS_herb = round(cbind(TSS_herb[, 1:4] / sum(TSS_herb[, 1:4]) * 100, TSS = TSS_herb[, 5]), 2)
    )
  )

colnames(tss_full)[2:5] <- c("a = true positive", "b = false positive", "c = false negative", "d = true negative")

tss_full

save(tss_full, file = "./output/hybrid_tss_table.RData")
```

# create different objects
```{r}
T_pred <- quantile(predator_interactions$pred1, c(.75))
T_omni <- quantile(omnivores_interactions$pred1, c(.75))
T_herb <- quantile(herbivore_interactions$pred1, c(.75))

############################################################################################################################################
### Rule-based priority AND RF>0.2
############################################################################################################################################

# rule based interactions
hybrid_rule_based_interactions <- hybrid_interact_con_res_ib

names(hybrid_rule_based_interactions)[names(hybrid_rule_based_interactions) == "value"] <- "weight"

hybrid_rule_based_edges <-
  hybrid_rule_based_interactions %>%
  dplyr::select(
    con_code, res_code, weight
  )

#rename and change the order
names(hybrid_rule_based_edges)[names(hybrid_rule_based_edges) == "con_code"] <- "from"
names(hybrid_rule_based_edges)[names(hybrid_rule_based_edges) == "res_code"] <- "to"
hybrid_rule_based_graph <- prepare_graphs_plotting(hybrid_rule_based_edges, hybrid_data_trait_phylo, .column_name = "unique_id")

############################################################################################################################################
## Rule-based + randomForest priority when RF>0.2 priority
############################################################################################################################################

hybrid_rf_priority_interactions <- hybrid_interact_con_res_ib

### select predator interactions
  
hybrid_rf_priority_interactions[
    hybrid_rf_priority_interactions$con_feeding_type.predator == 1 &
      hybrid_rf_priority_interactions$value == 1 &
      hybrid_rf_priority_interactions$pred1 < T_pred,]$value <- 0

hybrid_rf_priority_interactions[
    hybrid_rf_priority_interactions$con_feeding_type.predator == 1 &
      hybrid_rf_priority_interactions$value == 0 &
      hybrid_rf_priority_interactions$pred1 >= T_pred,]$value <- 1

### select omnivore interactions
 if(length(
   hybrid_rf_priority_interactions[
     hybrid_rf_priority_interactions$con_feeding_type.omnivore == 1 &
     hybrid_rf_priority_interactions$value == 1 &
     hybrid_rf_priority_interactions$pred1 < T_omni,]$value
   )>0) {
   hybrid_rf_priority_interactions[
     hybrid_rf_priority_interactions$con_feeding_type.omnivore == 1 &
       hybrid_rf_priority_interactions$value == 1 &
       hybrid_rf_priority_interactions$pred1 < T_omni,]$value <- 0
   }
  
if(length(
    hybrid_rf_priority_interactions[
      hybrid_rf_priority_interactions$con_feeding_type.omnivore == 1 &
      hybrid_rf_priority_interactions$value == 0 &
      hybrid_rf_priority_interactions$pred1 >= T_omni,]$value
    )>0) {
    hybrid_rf_priority_interactions[
      hybrid_rf_priority_interactions$con_feeding_type.omnivore == 1 &
        hybrid_rf_priority_interactions$value == 0 &
        hybrid_rf_priority_interactions$pred1 >= T_omni,]$value <- 1
}

### select herbivore interactions
hybrid_rf_priority_interactions[
hybrid_rf_priority_interactions$con_feeding_type.herbivore == 1 &
  hybrid_rf_priority_interactions$value == 1 &
  hybrid_rf_priority_interactions$pred1 < T_herb,]$value <- 0

hybrid_rf_priority_interactions[
  hybrid_rf_priority_interactions$con_feeding_type.herbivore == 1 &
    hybrid_rf_priority_interactions$value == 0 &
    hybrid_rf_priority_interactions$pred1 >= T_herb,]$value <- 1

colnames(hybrid_rf_priority_interactions)[colnames(hybrid_rf_priority_interactions)=="value"] <- "weight"

hybrid_rf_priority_edges <-
  hybrid_rf_priority_interactions %>%
  dplyr::select(
    con_code, res_code, weight
  )

#rename and change the order
names(hybrid_rf_priority_edges)[names(hybrid_rf_priority_edges) == "con_code"] <- "from"
names(hybrid_rf_priority_edges)[names(hybrid_rf_priority_edges) == "res_code"] <- "to"
hybrid_rf_priority_graph <- prepare_graphs_plotting(hybrid_rf_priority_edges, hybrid_data_trait_phylo, .column_name = "unique_id")
```


```{r}
############################################################################################################################################
## RandomForest only when RF>0.2 priority
############################################################################################################################################

hybrid_rf_interactions <- hybrid_interact_con_res_ib

### select predator interactions
hybrid_rf_interactions[hybrid_rf_interactions$con_feeding_type.predator==1 & hybrid_rf_interactions$pred1 >= T_pred,]$value <- 1
hybrid_rf_interactions[hybrid_rf_interactions$con_feeding_type.predator==1 & hybrid_rf_interactions$pred1 < T_pred,]$value <- 0

### select omnivore interactions
hybrid_rf_interactions[hybrid_rf_interactions$con_feeding_type.omnivore==1 & hybrid_rf_interactions$pred1 >= T_omni,]$value <- 1
hybrid_rf_interactions[hybrid_rf_interactions$con_feeding_type.omnivore==1 & hybrid_rf_interactions$pred1 < T_omni,]$value <- 0
  
### select herbivore interactions
hybrid_rf_interactions[hybrid_rf_interactions$con_feeding_type.herbivore==1 & hybrid_rf_interactions$pred1 >= T_herb,]$value <- 1
hybrid_rf_interactions[hybrid_rf_interactions$con_feeding_type.herbivore==1 & hybrid_rf_interactions$pred1 < T_herb,]$value <- 0

colnames(hybrid_rf_interactions)[colnames(hybrid_rf_interactions)=="value"] <- "weight"

hybrid_rf_edges <-
  hybrid_rf_interactions %>%
  dplyr::select(
    con_code, res_code, weight
  )
#rename and change the order
names(hybrid_rf_edges)[names(hybrid_rf_edges) == "con_code"] <- "from"
names(hybrid_rf_edges)[names(hybrid_rf_edges) == "res_code"] <- "to"
hybrid_rf_graph <- prepare_graphs_plotting(hybrid_rf_edges, hybrid_data_trait_phylo, .column_name = "unique_id")

# add detritus
hybrid_rule_based_graph <- add_detritus_to_graph(.graph = hybrid_rule_based_graph)
hybrid_rf_priority_graph <- add_detritus_to_graph(.graph = hybrid_rf_priority_graph)
hybrid_rf_graph <- add_detritus_to_graph(.graph = hybrid_rf_graph)

# save predictions
# rules based only
save(hybrid_rule_based_interactions, file = "./output/hybrid_rule_based_interactions.RData")
# RF priority
save(hybrid_rf_priority_interactions, file = "./output/hybrid_rf_priority_interactions.RData")
# RF only
save(hybrid_rf_interactions, file = "./output/hybrid_rf_interactions.RData")

# save prediction objects with different consensus filters
# rules based only
save(hybrid_rule_based_edges, file = "./output/hybrid_rule_based_edges.RData")
# RF priority
save(hybrid_rf_priority_edges, file = "./output/hybrid_rf_priority_edges.RData")
# RF only
save(hybrid_rf_edges, file = "./output/hybrid_rf_edges.RData")

# save prediction objects with different consensus filters
# rules based only
save(hybrid_rule_based_graph, file = "./output/hybrid_rule_based_graph.RData")
# RF priority
save(hybrid_rf_priority_graph, file = "./output/hybrid_rf_priority_graph.RData")
# RF only
save(hybrid_rf_graph, file = "./output/hybrid_rf_graph.RData")
```